---
title: "Exam_Notes"
author: "Elwood Fong"
date: "2025-04-13"
output: 
  rmdformats::downcute:
    number_sections: false
    toc_depth: 3
    use_bookdown: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library Initialization

```{r, message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggalt)
library(naniar)
```

## Data Initialization

```{r}
data <- iris
# data <- read.csv("FilePath\to\csvFile")

# Artificially Add Column
data$Size <- ifelse(data$Sepal.Length > 5, "Large", "Small")
```

## Data Cleaning

### Convert "character" to factor (Turns it Categorical)

```{r}
data$Size <- as.factor(data$Size)
```

### Visualizing Missingness of Data
```{r}
vis_miss(data)
```

### Remove NA Values

```{r}
data <- na.omit(data)
```

### Getting Dimensions of Data (See Observations and Variables)

```{r}
dim(data)
# Outputs: 
# Observations: 150    Variables: 6
```

### Changing Data Dimensions

```{r}
grouped_data <- data
grouped_data %>%
  group_by(Species) %>%
  summarise(
    mean_Petal.Length = mean(Petal.Length, na.rm = TRUE), 
    median_Petal.Length = median(Petal.Length, na.rm = TRUE))
```

```{r}
grouped_data <- data
grouped_data %>%
  group_by(Species) %>%
  summarise(
    mean_Petal.Length = mean(Petal.Length, na.rm = TRUE), 
    median_Petal.Length = median(Petal.Length, na.rm = TRUE)) %>%
  pivot_longer(cols=c("mean_Petal.Length", "median_Petal.Length"), names_to="summary", values_to="length")
```

### Scaling Data {#datascaling}

```{r}
ScaledData <- data %>%
  dplyr::select(-Species, -Size) %>%
  scale()
head(ScaledData)
# Output:
# Table of Scaled Data
# Should only be done with numerical variables
```

## Reading Data

### Get Row with Minimum Petal.Length

```{r}
library(tidyr)
data %>%
  filter(data$Petal.Length==min(data$Petal.Length))
```

### Count Number of NA Values

```{r}
sum(is.na(data))
# Outputs:
# Number of NA values: 0
```

### Count Number of rows with NA Values

```{r}
sum(!complete.cases(data))
# Outputs:
# Number of rows with NA values: 0
```

### Check Variable Data Types
```{r}
sapply(data,class)
# Outputs:
# Sepal.Length  Sepal.Width Petal.Length  Petal.Width       Species         Size 
#    "numeric"    "numeric"    "numeric"    "numeric"      "factor"      "factor"

# Type of Variable:
# Sepal.Length  Sepal.Width Petal.Length  Petal.Width       Species         Size 
#  "numerical"  "numerical"  "numerical"  "numerical" "categorical" "categorical"
```

### Constructing a Contingency Table
```{r}
contingency_table <- table(data$Species, data$Size)
contingency_table
```

## Analyzing Data

### Chi-Squared Test

#### Requirements
- Comparing 2 CATEGORICAL variables
- H~0~  = Null Hypothesis 
  - There is no relationship between variables
  - TRUE only if `p-value` > 0.05
- H~a~  = Alternate Hypothesis 
  - There is a relationship between variables
  - TRUE only if `p-value` < 0.05

#### Application
```{r}
chisq_tab <- table(data$Species, data$Size)
chisq_result = chisq.test(chisq_tab)
chisq_result
```

#### Chi-Squared Test Result Components

| R Code                     | Meaning / Description                                                     | Interpretation / Usage                         |
|---------------------------|----------------------------------------------------------------------------|------------------------------------------------|
| `chisq_result$statistic`  | Chi-squared Test Statistic (X²)                                            | Higher value → More difference between observed and expected counts and stronger evidence against H₀   |
| `chisq_result$parameter`  | Degrees of Freedom (df)                                                    | Depends on table size: (rows-1) × (cols-1) → just tells you how many categories you're comparing   |
| `chisq_result$p.value`    | P-value (significance level)                                               | Small p-value (< 0.05) → Reject H₀           |
| `chisq_result$observed`   | Observed counts from your data                                             | Actual values in your data table              |
| `chisq_result$expected`   | Expected counts under the null hypothesis                                  | What values you'd expect if no association, if Expected Count ≥ 5 → The Chi-squared test is reliable.    |
| `chisq_result$residuals`  | Raw Residuals = (Observed - Expected) / sqrt(Expected)                    | Measures difference between observed and expected → if bigger, then stronger evidence against H₀        |
| `chisq_result$stdres`     | Standardized Residuals (scaled residuals) →                               | Values Between -2 and 2	supports H₀, Bigger than 2 or less than -2 supports H~a~ |

---

```{r}
chisq_result$statistic       # Test statistic
chisq_result$parameter       # Degrees of freedom
chisq_result$p.value         # P-value
chisq_result$observed        # Observed table
chisq_result$expected        # Expected counts
chisq_result$residuals       # Residuals
chisq_result$stdres          # Standardized residuals
```

#### Chi-Squared Test Result Analysis
```{r}
chisq_result$stdres > 2 | chisq_result$stdres < -2
```

### K Means Clustering

> Requires Data Scaling from [data scaling](#datascaling)

```{r}
set.seed(256)

k_means_data <- data

kmeans_result <- kmeans(ScaledData, centers = 2)

k_means_data$cluster <- as.factor(kmeans_result$cluster)

ggplot(k_means_data, aes(x = Petal.Width, y = Petal.Length, color = cluster)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_encircle(aes(group = cluster), expand = 0.05, spread = 0.05,
                color = "black", linetype = "dashed", size = 1.2) +
  labs(title = "Title",
       x = "Petal.Width", y = "Petal.Length")
```

### ANOVA

#### Requirements
- Comparing 1 CATEGORICAL and 1 NUMERICAL variable
- H~0~  = Null Hypothesis 
  - There is no relationship between variables
  - TRUE only if `p-value` > 0.05
- H~a~  = Alternate Hypothesis 
  - There is a relationship between variables
  - TRUE only if `p-value` < 0.05
  
#### ANOVA Result Componenets

| ANOVA Output Column | Meaning / Description                                                | Interpretation / Usage                                    |
|--------------------|----------------------------------------------------------------------|-----------------------------------------------------------|
| `Df`               | Degrees of Freedom                                                   | Number of groups - 1 for factor; Number of observations - groups for residuals |
| `Sum Sq`           | Sum of Squares                                                       | Total variation explained by factor or left in residuals  |
| `Mean Sq`          | Mean Square = Sum Sq / Df                                            | Average variation per degree of freedom                  |
| `F value`          | F-statistic = Mean Sq of Factor / Mean Sq of Residuals               | Higher value → Stronger evidence against H₀ (factor affects response variable) |
| `Pr(>F)`           | P-value for F-statistic                                              | Small p-value (< 0.05) → Reject H₀ → Factor has significant effect |
  
#### 1 way ANOVA

```{r}
anova_result = aov(Petal.Width~Species, data)
summary(anova_result)
```

```{r}
TukeyHSD(anova_result)
```

#### 2 way ANOVA
```{r}
aov_2_result <- aov(Petal.Width ~ Species + Size, data)
summary(aov_2_result)
```


```{r}
TukeyHSD(aov_2_result)
```