---
title: "Sample_Rmd"
output: bookdown::html_document2
date: "2025-02-14"
author: "Author"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# This is the config for this Rmd file setting the default echo to TRUE
```

### Contents
- [Abstract](#abstract)
- [Introduction](#introduction)
- [Results](#results)
- [Discussion](#discussion)
- [Conclusion](#conclusion)

### Abstract {#abstract}
No conclusion could be drawn due to the unreliability of the data. 
However, the conclusion drawn from the data was a positive relationship between caffeine intake and sleep quality for males and the inverse for females and other. 


### Introduction {#introduction}
This is a sample report (or a template for a report) using data taken from [kaggle.com](https://www.kaggle.com/datasets/arsalanjamal002/student-sleep-patterns). 

The intention of this report is to determine the relationship between Caffeine Intake and Sleep Quality and Gender.

### Results {#results}

```{r, echo = FALSE}
student_sleep_patterns <- read.csv("~/Desktop/University/AMED3002/student_sleep_patterns.csv", header=T)
library(ggplot2)
```

``` {r caffeine-sleep-plot, fig.cap = "A Jitter Plot that explores the correlation between Caffeine Intake and Sleep Quality by gender.", echo = TRUE}

p <- ggplot(data = student_sleep_patterns, aes(x = Caffeine_Intake, y = Sleep_Quality, col = Gender))

p + labs(title = "Relationship Between Caffeine Intake and Sleep Quality by Gender", subtitle = "Data Collected from Kaggle, Showing Caffeine Intake and Sleep Quality Patterns Across Genders", caption = "*Each point represents an individual, with color representing gender.") +
  geom_jitter(width = 0.2, height = 0.2, size = 2, alpha = 0.5) +
  stat_smooth(method ="lm", formula = y ~ poly(x,2), aes(fill = Gender), alpha = 0.2, se = TRUE)
```

### Discussion {#discussion}
From Figure \@ref(fig:caffeine-sleep-plot) it can be observed that:

- In females and other:
  - A higher caffeine intake can lead to a lower sleep quality 
  - A lower caffeine intake can lead to a higher sleep quality
- In males:
  - A higher caffeine intake can lead to a higher sleep quality
  - A lower caffeine intake can lead to a lower sleep quality

However, this data is only limited to university students with the age range of 18 to 24.
Additionally, it is unknown how this data was collected. 

### Conclusion {#conclusion}
In conclusion, due to the unreliability of the data (in not knowing how the data was collected) no conclusions can be drawn.
